## missing value

```{r}
dataset= na.omit(dataset)
```

We analyzed the database to identify missing values using the provided code, and we found missing values that we have deleted. As a result, the number of rows decreased from 846 to 816.

## outliers analysis

radius_ratio founded outliers 3

```{r}
quartiles <- quantile(dataset$radius_ratio, probs=c(.25, .75), na.rm = FALSE)
IQR <- IQR(dataset$radius_ratio)

Lower <- quartiles[1] - 1.5*IQR
Upper <- quartiles[2] + 1.5*IQR 


data_no_outlier <- subset(dataset, dataset$radius_ratio > Lower & dataset$radius_ratio < Upper)
dim(data_no_outlier)
```

pr.axis_aspect_ratio outliers 8

```{r}
quartiles <- quantile(dataset$pr.axis_aspect_ratio, probs=c(.25, .75), na.rm = FALSE)
IQR <- IQR(dataset$pr.axis_aspect_ratio)

Lower <- quartiles[1] - 1.5*IQR
Upper <- quartiles[2] + 1.5*IQR 


data_no_outlier <- subset(dataset, dataset$pr.axis_aspect_ratio > Lower & dataset$pr.axis_aspect_ratio < Upper)
dim(data_no_outlier)
```

max.length_aspect_ratio outliers is 5

```{r}
quartiles <- quantile(dataset$max.length_aspect_ratio, probs=c(.25, .75), na.rm = FALSE)
IQR <- IQR(dataset$max.length_aspect_ratio)

Lower <- quartiles[1] - 1.5*IQR
Upper <- quartiles[2] + 1.5*IQR 


data_no_outlier <- subset(dataset, dataset$max.length_aspect_ratio > Lower & dataset$max.length_aspect_ratio < Upper)
dim(data_no_outlier)
    

```

scaled_variance outliers is 1

```{r}
quartiles <- quantile(dataset$scaled_variance, probs=c(.25, .75), na.rm = FALSE)
IQR <- IQR(dataset$scaled_variance)

Lower <- quartiles[1] - 1.5*IQR
Upper <- quartiles[2] + 1.5*IQR 


data_no_outlier <- subset(dataset, dataset$scaled_variance > Lower & dataset$scaled_variance < Upper)
dim(data_no_outlier)
```

scaled_variance.1 outliers is 2

```{r}
quartiles <- quantile(dataset$scaled_variance.1, probs=c(.25, .75), na.rm = FALSE)
IQR <- IQR(dataset$scaled_variance.1)

Lower <- quartiles[1] - 1.5*IQR
Upper <- quartiles[2] + 1.5*IQR 


data_no_outlier <- subset(dataset, dataset$scaled_variance.1 > Lower & dataset$scaled_variance.1 < Upper)
dim(data_no_outlier)
```

scaled_radius_of_gyration.1 is outliers is 10

```{r}
quartiles <- quantile(dataset$scaled_radius_of_gyration.1, probs=c(.25, .75), na.rm = FALSE)
IQR <- IQR(dataset$scaled_variance.1)

Lower <- quartiles[1] - 1.5*IQR
Upper <- quartiles[2] + 1.5*IQR 


data_no_outlier <- subset(dataset, dataset$scaled_radius_of_gyration.1 > Lower & dataset$scaled_radius_of_gyration.1 < Upper)
dim(data_no_outlier)
```

skewness_about is outliers is 3

```{r}
quartiles <- quantile(dataset$skewness_about, probs=c(.25, .75), na.rm = FALSE)
IQR <- IQR(dataset$skewness_about)

Lower <- quartiles[1] - 1.5*IQR
Upper <- quartiles[2] + 1.5*IQR 


data_no_outlier <- subset(dataset, dataset$skewness_about > Lower & dataset$skewness_about < Upper)
dim(data_no_outlier)
```

skewness_about.1 is outliers is 3

```{r}

quartiles <- quantile(dataset$skewness_about.1, probs=c(.25, .75), na.rm = FALSE)
IQR <- IQR(dataset$skewness_about.1)

Lower <- quartiles[1] - 1.5*IQR
Upper <- quartiles[2] + 1.5*IQR 


data_no_outlier <- subset(dataset, dataset$skewness_about.1 > Lower & dataset$skewness_about.1 < Upper)
dim(data_no_outlier)
    
```
