---
title: "Vehicle Silhouettes"
output: html_notebook
---
---
title: "Vehicle Silhouettes"

Goal of collecting this dataset:
The goal of collecting this dataset is to determine the category of a given vehicle silhouette among three types (car, van, bus) by utilizing a set of extracted silhouette-based features.

Classification:
The classification of vehicle silhouettes involves the utilization of a set of distinctive attributes extracted from these silhouettes. The primary objective is to categorize these silhouettes into one of three predefined classes: cars, vans, or buses.

Defect prediction:
Defect prediction in a dataset of vehicle silhouettes involves spotting potential issues or irregularities within these silhouette images. This process may encompass the detection of inconsistencies or abnormalities in the silhouettes, including the presence of missing data points, which can be problematic. Leveraging data mining techniques, defect prediction enhances the dataset's overall quality and dependability, ensuring it faithfully reflects the desired information. This, in turn, reduces the risk of producing inaccurate or deceptive outcomes when the dataset is applied in subsequent tasks and applications, such as object recognition and classification.

Link Of dataset:
https://www.kaggle.com/datasets/pritech/vehicle-silhouettes/data

Number of attribute:
19
Number of objects:
847

```{r}
"The following code is to show raw data of our dataset and type of attribute"
##read file named "vehicle.csv" and save it in dataframe named "dataset"
dataset=read.csv("C:/Users/Tshah/OneDrive/Desktop/326/project/vehicle.csv")
View(dataset)

##Structure of our dataset
str(dataset)
"Lets take a look at our variable names and sizes:"

##Number of rows:
nrow(dataset)

#Number of column:
ncol(dataset)

##Names:
names(dataset)
```




---
The missing values of our dataset
```{r}

# Finding the missing values:
sum(is.na(dataset))

```



---
---
Plotting methods

We have identified eight crucial features in our dataset that will serve as the foundation for our classification task and play a pivotal role in effectively categorizing our data.
The following bullet points outline the primary eight features we've selected and provide the rationale for choosing each one.

1)Compactness is a valuable feature because it offers insights into the general shape of a car, which can be a useful indicator of the car's type.

2)Circularity is important because it helps us assess how closely a car's shape resembles that of a perfect circle, and each car type exhibits unique circularity traits.

3)The inclusion of "max.length.rectangularity" is justified by its representation of the maximum length of the car's bounding rectangle. This attribute is significant for distinguishing cars based on their overall size and proportions.

4)The rationale behind selecting "Pr.axis_aspect_ratio" is to convey information about the elongation or stretching of a car's shape. This is important because different car types can display diverse aspect ratios, and this feature helps capture such variations.

5)The inclusion of "Scaled variance" is driven by its capacity to offer insights into the dispersion of data points along the outline of the vehicle. Variations in this spread can serve as indications of distinct vehicle types.

6)The inclusion of the "skewness_about" feature is based on its ability to convey information regarding the asymmetry present in the shape of vehicles. This aspect of asymmetry tends to differ among various vehicle types, including cars, vans, and buses.

7)Hollows_ratio" is chosen because it reflects the degree of concavity or hollowness in the shape of a vehicle, and this attribute is valuable for aiding in the classification process.

In the following code we will apply different type of of plotting methods to understand dataset
```{r}
install.packages("dplyr")
library(dplyr)
     dataset$class %>% table() %>% barplot()
     ## As we can see from barplot car count is more than bus and van
install.packages("scatterplot3d") # install only one time then put this command as comment after installation
library(scatterplot3d)
scatterplot3d(dataset$compactness, dataset$hollows_ratio, dataset$pr.axis_aspect_ratio)

dist.matrix <- as.matrix(dist(dataset[, 1:9]))
heatmap(dist.matrix) 

```